Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2022-09-24 response-reviewers-2.tex

\hspace*{8cm} Yours,  \\

\centerline{\includegraphics[width=2.5cm,height=2.5cm]{jj_sign.eps}}

\hspace*{8cm} Juan J. Merelo\\
\hspace*{8cm} Catedrático de Universidad (Full professor)

%%%% Rev 3

{\it

{\bf Reviewer \#3}: Paper Summary:
This study surveyed teachers' feedback about their opinions regarding introducing the messaging platforms and chatbots in their classes, understand their needs, and gathering information about the various educational use cases where these tools are valuable. There are some suggestions below to improve the research work.
1.      Does the article address a topic of current concern about the aims of the journal and to what you perceive to be the current state of ICT concerning education? The article addresses issues bothering the introduction of messaging platforms and chatbots to augment teaching and learning, which I consider important concerning the current state of ICT education.}

% -> answer 1

The authors thank the comment for noticing the relevance of the topic and how it suits the journal scope.

{\it 2.      Does the abstract give a clear account of the scope of the paper? The abstract gave an overview of the paper. However, the key findings of the paper were not reported in the abstract. I suggest the authors summarize the study's findings and report them briefly in the abstract.}

% -> answer 2

The abstract as well as other sections in the paper have been rewritten. The abstract now summarizes the study's findings and reports them briefly.

% -> answer 3

{\it 3.      Do the keywords adequately reflect the paper? The authors should consider replacing Tutorship with teachers.}

After consideration, the authors agree with the comment and we have replaced teachers for Tutorship in the keywords.

% -> answer 4

{\it 4.      If the article is concerned with research activity, has a sound methodology been utilized and described? A sound methodology was utilized and described. However, it was observed that prior studies have not previously validated the measurement instrument used in collecting data. Thus, the authors need to conduct reliability and validity tests for their proposed measurement instrument used in collecting data.}

TO DO

% -> answer 5,6,7,8

{\it 5.      Does the article distinguish clearly between opinion and empirical evidence? Yes
6.      Does the article contribute to a critical understanding of the issues? Yes
7.      Does the article alert readers to significant new developments in the subject area? Yes
8.      Does the article take into account relevant contemporary literature in the area? Yes.}

The authors want to thank the reviewer for evaluating positively those aspect of the paper.

% -> answer 9

9.      Does the article clearly reference citations and quotations using the APA 7th edition? Partly, some of the references do not follow APA 7th edition. For example, the following references don't follow APA 7th edition referencing style
     a.   O'Boyle, 2014. O'Boyle, N. (2014). Front row friendships: Relational dialectics and identity negotiations by mature students at univerty. Communication Education, 63(3):169-191
     b. Panah and Babar, 2020. Panah, E. and Babar, M. Y. (2020). A survey of WhatsApp as a tool for instructor-learner dialogue, learner-content dialogue, and learner-learner dialogue. International Journal of Educational and Pedagogical
 Sciences, 14(12):1198 - 1205.
     c. Perez et al., 2020. Perez, J. Q., Daradoumis, T., and Puig, J. M. M. (2020). Rediscovering the use of chatbots in education: A systematic literature review. Computer Applications in Engineering Education, 28(6):1549- 1565.
     d. Ranoliya et al., 2017. Ranoliya, B. R., Raghuwanshi, N., and Singh, S. (2017). Chatbot for university related faqs. In 2017 International Conference on Advances in Computing, Communications and Informatics (ICACCI), pages 1525-1530. IEEE.
     e. Roblyer et al., 2010. Roblyer, M., McDaniel, M., Webb, M., Herman, J., and Witty, J. V. (2010). Findings on facebook in higher education: A comparison of college faculty and student uses and perceptions of social networking sites. The Internet and Higher Education, 13(3):134 - 140.

% https://www.scribbr.com/apa-style/apa-seventh-edition-changes/

% -> answer 10      
{\it 10.     Is the article written in a clear and intelligible style to a reasonably well-informed international professional readership? Yes. However, there are a few grammatical errors. I suggest the authors proofread the manuscript to correct them.}

The authors apologise for the grammatical errors and have strived to fix them and to avoid them in the new paragraphs.

% -> answer 11
{\it 11.     Other few comments the authors should consider:
   a.   Reorganize the conclusion section to highlight the main aim of the research work briefly, key findings and the significance of the work, and future research work.
   b.   Some of the paragraphs in the conclusion section can be moved to Discussion. The Discussion section can be captioned as Discussion and Implications to cater for some of the paragraphs that will be moved to this section. For example, paragraphs on Page 23, Line 16, "In our survey…" and Page 23, Line 37, "One of the questions we made in the second survey…" should be moved to the Discussion and Implication section.}

To address these final comments:
    a. as cited before, the abstrac and other sections have been rewritten highlighting the key findings and the significance of the work.

    b. TO DO

%%%%%% Rev 4

{\bf Reviewer \#4}: {\it First of all, this paper seems to be the main topic of 'Analysis from the teacher's perspective of chatbots and messaging platforms in class'.
It seems that simple survey results were analyzed and grouped.
It seems like a lot of effort was put into it, but I doubt whether it has any academic significance.
Why did you choose 4 questions? Is it appropriate for a chatbot and messaging platform if I just ask and group this? I am not sure.
Looking at the research results and conclusions, there are some contents that are difficult to see as valid because the evidence for the researcher's conclusions is so abrupt and abstract.}

### 2022-09-24 edubots_surveys.Rnw
Update edubots_surveys.Rnw
Introduction completely rewritten

\textcolor{blue}{\textit{chatbots} are autonomous agents able to communicate and interact with humans (or other chatbots) in a huma-like way \cite{Gong2008}, answering question and even posing their own ones to the user. 
Thus, we can say that they are close to passing a Turing Test \cite{moor2003turing}. They are implemented by means of different Artificial Intelligence techniques, such as natural language processing.
Inside the plethora of possible applications of chatbots, education is one of the most prolific nowadays \cite{clarizia2018,Smutny2020}. In this domain, they facilitate the so-called 
\textit{personalized learning}, i.e. they adapt themselves to students’ pace of learning and provide online tutoring outside the classroom, which have been proved to be very effective increasing their 
engagement and participation in the subjects \cite{agarwal2020}. Moreover chatbots' tutoring possibilities aid to reduce the stress of students \cite{Daniel2016,agarwal2020} by avoiding the face-to-face tutoring with their teacher, 
or asking questions in front of the class. 
}

This is why many students resort to silence in the classroom and tend to
contact their teachers later via email.  Consequently, this can lead to not only
a delay in obtaining answers to their questions but can also cause a significant
increase in their teachers’ workload, especially when factoring in the
disproportion between the class size and the number of teachers. Hence, the
chatbot technology has a potential to mitigate this problem by providing answers
to students’ questions and facilitating a dynamic and autonomous learning
experience \cite{Griol2014,kim2020}. In addition, using an automated system such
as a chatbot, can draw teachers’ attention to topics that students struggle to
comprehend or need further assistance in understanding.

Given the potential benefits that chatbots could bring to the classroom, the
focus of the
EDUBOTS KA2 European project \footnote{https://www.edubots.eu},
the sponsor of this study, is to
\textcolor{black}{explore best practices and innovative use of chatbots, and to create 
a learning community of educators in higher education institutions.
The majority of studies conducted so far have studied how these 
messaging applications and/or chatbots are used to deliver personalised
learning in classrooms that occurs anytime anywhere,
promote collaborative learning experiences, group discussions \cite{panah2020study}
and boost students' sense of belonging to their institutions
\cite{abbas2021onlinechat}. Hence, due to the paucity of research
that explores the use of these technologies from the teachers' perspectives,
this paper aims to investigate their opinions/needs and the
challenges/opportunities of adopting these technologies in classrooms
to paint a better picture of how they can positively contribute to enhancing
the learning process in higher education institutions.}

\textcolor{black}{In this study, data collection involved two phases:}
 \begin{itemize}
\item The first phase aimed to collect feedback from students about their actual
  information about students’ preferences of tools/applications for chatting and
  messaging, how they use them in educational context, who do they like to be
  with in the class messaging groups, and their expectations of the chatbots in
  terms of assisting them during their learning process. Therefore, two surveys
  for bachelor and master degree students at the University of Granada (Spain)
  were designed and answers from more than 250 students were collected. The key
  findings of the students' surveys reveal that they prefer to use messaging
  applications, such as Telegram or WhatsApp as they are familiar to
  them. Students also expressed interest in chatbots that can assist them with
  organizing their course schedules, help them access their assignment grades
  and facilitate searching for course's resources.
\item The results of the first phase were fed into the second phase of the
  study, which focused on teachers’ opinions about using chatbots and messaging
  applications in classes. This is what is going to be presented in this paper.
\end{itemize}

@@

\begin{itemize}
\item RQ1 - Are teachers already using messaging apps in their classes?
\item RQ2 - Which chatbots' features would teachers find useful in their
  classes?
\item RQ3 - Which kind of interaction do teachers prefer with their students?
\item RQ4 - What kind of interaction media features do teachers value the most?
\end{itemize}

We will try to answer these questions through the analysis of questions and
answers in the two surveys. Eventually, what we want to obtain is a series of
recommendations to make a successful deployment of chatbot technologies (and, in
some cases, general instant messaging applications) in higher education\textcolor{black}{; from the response to RQ1 and RQ4 we will try to recommend specific technologies or applications to be used in the classroom; the response to RQ2 and RQ3 will help us recommend chatbots features or specific platforms; and finally, from RQ3 we will also try to find best practices in the adoption of messaging platforms and they accompanying chatbots.}

>>>>

\textcolor{blue}{\textit{Chatbots} are autonomous agents able to communicate and interact with humans (or other chatbots) in a huma-like way \cite{Gong2008}, answering question and even posing their own ones to the user. 
Thus, we can say that they are close to passing a Turing Test \cite{moor2003turing}. They are implemented by means of different Artificial Intelligence techniques, such as natural language processing.
Inside the plethora of possible applications of chatbots, education is one of the most prolific nowadays \cite{clarizia2018,Smutny2020}. In this domain, they facilitate the so-called 
\textit{personalized learning}, i.e. they adapt themselves to students’ pace of learning and provide online tutoring outside the classroom, which have been proved to be very effective increasing their 
engagement and participation in the subjects \cite{agarwal2020}. Moreover chatbots' tutoring possibilities aid to reduce the stress of students \cite{Daniel2016,agarwal2020} by avoiding the face-to-face tutoring with their teacher, 
or asking questions in front of the class. This situation normally conducts to a silence of the student and a later contact with the teacher by e-mail, which, in turn, will mean a workload for the teacher, that could be very 
significant given the existing disproportion between number of students per each teacher.  
}
Hence, the chatbot technology has a potential to mitigate this problem by providing answers to students’ questions and facilitating a dynamic and autonomous learning
experience \cite{Griol2014,kim2020}. In addition, using an automated system such as a chatbot, can draw teachers’ attention to topics that students struggle to comprehend or need further assistance in understanding.

Given the potential benefits that chatbots could bring to the classroom, the
focus of the EDUBOTS KA2 European project \footnote{https://www.edubots.eu},
the main sponsor of this study, is to explore best practices and innovative use of chatbots, 
and to create a learning community of educators in higher education institutions.
\textcolor{blue}{Researchers and teachers participating in this project have developed several studies. Most of these works have been focused on how the 
messaging applications and/or chatbots are normally applied to deliver personalised learning in classrooms that occurs anytime anywhere,
promote collaborative learning experiences, group discussions \cite{panah2020study} and boost students' sense of belonging to their institutions
\cite{abbas2021onlinechat}.}
\textcolor{blue}{The majority of these studies were conducted focusing on the teachers' perspectives, so this paper aims to provide more conclusions regarding 
their opinions and needs, as well as the existing challenges or opportunities that arise when adopting these technologies in classrooms. This will lead to clarify
how they can positively contribute to enhancing the learning process in higher education institutions.}

In this study, data collection involved two phases:
 \begin{itemize}
\item The first phase aimed to collect feedback from students about their actual
  information about students’ preferences of tools/applications for chatting and
  messaging, how they use them in educational context, who do they like to be
  with in the class messaging groups, and their expectations of the chatbots in
  terms of assisting them during their learning process. Therefore, two surveys
  for bachelor and master degree students at the University of Granada (Spain)
  were designed and answers from more than 250 students were collected. The key
  findings of the students' surveys reveal that they prefer to use messaging
  applications, such as Telegram or WhatsApp as they are familiar to
  them. Students also expressed interest in chatbots that can assist them with
  organizing their course schedules, help them access their assignment grades
  and facilitate searching for course's resources.
\item The results of the first phase were fed into the second phase of the
  study, which focused on teachers’ opinions about using chatbots and messaging
  applications in classes. This is what is going to be presented in this paper.
\end{itemize}

@@

\begin{itemize}
\item RQ1 - Do teachers use now or want to use messaging apps in their classes?
\item RQ2 - Which chatbots' features would teachers find useful for this purpose?
\item RQ3 - Which kind of interaction do teachers prefer with their students?
\item RQ4 - What kind of interaction media features do teachers value the most?
\end{itemize}

We will try to answer these questions through the analysis of questions and
answers in the two aforementioned surveys. Eventually, what we want to obtain is a series of
recommendations to make a successful deployment of chatbot technologies (and, in
some cases, general instant messaging applications) in higher education\textcolor{black}{; from the response to RQ1 and RQ4 we will try to recommend specific technologies or applications to be used in the classroom; the response to RQ2 and RQ3 will help us recommend chatbots features or specific platforms; and finally, from RQ3 we will also try to find best practices in the adoption of messaging platforms and they accompanying chatbots.}
